### 2020-10-16 doc/manuscript/Dow_manuscript.Rmd
Includes edits from KAT

---
title: ''
output:
  word_document:
    fig_caption: yes
  pdf_document: default
bibliography: references.bib
csl: global-change-biology.csl
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

The significance of wood type in regards to wood and leaf phenology likely implies evolutionary differences and adaptations to local climates where these trees can be found. Ring-porous trees have large xylem cells which are prone to embolism during the winter. These cells are repaired before leaf development, explaining why the 25% yearly growth milestone is reached around the same time as bud burst. Diffuse-porous xylem cells are smaller, and less prone to embolism during the Winter, meaning they do not need to be repaired in order to restore hydraulic functionality before beginning the leaf development process. Results from the paper by D’Orangeville et al. (in review) have shown that ring-porous trees reach 25 and 50% of their total yearly growth during Spring, while diffuse-porous trees reach these milestones in the Summer, but both reached 75% total annual growth around the same time. --Consequently, the Summer focused diffuse-porous trees are exposed to higher temperatures and drought risk than ring-porous trees.-- If leaf development is sped up when Spring temperatures are warmer, we should seek to understand whether this relationship holds, and how it potentially varies among trees of different wood-types.

@@

Here, I’ll insert a reference to Sean’s paper [@mcmahon_general_2015]. This is pulled from references.bib.

@@

@@

### Figures

**Figure 1**

Based on D'Orangeville Figure 1

```{r fig1, out.width="100%", echo=FALSE, fig.align="center", fig.cap= "Logistic growth model"}
# Created in RScripts/additional_figures.R
knitr::include_graphics("tables_figures/fig1.png")
```

**Figure 2**

Based on D'Orangeville Figure 3.a), but with relative growth

```{r fig2, out.width="100%", echo=FALSE, fig.align="center", fig.cap= "Percent total (modeled) growth"}
# Created in RScripts/additional_figures.R
knitr::include_graphics("tables_figures/fig2.png")
```

**Figure 3**

Similar to D'Orangeville Figure 4.a), but with cummalative relative growth

```{r fig3, out.width="100%", echo=FALSE, fig.align="center", fig.cap= "Cummulative percent total (modeled) growth"}
# Created in RScripts/additional_figures.R
knitr::include_graphics("tables_figures/fig3.png")
```

**Figure 4**

**Figure 5**

**Figure 6**

1. Model was fit to a random sample of 20 out of 106 tags to speed up computation.
1. Climwin mean temperatures are recentered at 65F to help sanity check intercepts and slopes. 65F was chosen because it is a "typical" climwinmean value
1. The multivariate model for $(DOY_{25}, DOY_{50}, DOY_{75})$ has:
    1. Fixed effects: The interaction between wood type (diffuse vs ring porous) and climwin mean
    1. Random effects: tag number (no species just yet)

```{r, echo = FALSE}
fig6_caption <- "Relationship between DOY and climwin mean tempeatures. For each 
observed climwin mean temperature value, the posterior mean of the fitted day of year
is represented by the solid blue line and credible intervals are represented by bands."
```

```{r fig6, out.width="100%", echo=FALSE, fig.align="center", fig.cap=fig6_caption}
# Created in RScripts/bert/wood_phenology_analysis_bert.R
knitr::include_graphics("tables_figures/fig6.png")
```

>>>>

---
title: ''
output:
  word_document:
    fig_caption: yes
bibliography: references.bib
csl: global-change-biology.csl
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

The significance of wood type in regards to wood and leaf phenology likely stems from evolutionary differences and adaptations to local climates where these trees can be found. Ring-porous trees have large xylem cells which are prone to embolism during the winter. These cells are repaired before leaf development, explaining why the 25% yearly growth milestone is reached around the same time as bud burst. Diffuse-porous xylem cells are smaller, and less prone to embolism during the Winter, meaning they do not need to be repaired in order to restore hydraulic functionality before beginning the leaf development process. Results from the paper by D’Orangeville et al. (in review) have shown that ring-porous trees reach 25 and 50% of their total yearly growth during Spring, while diffuse-porous trees reach these milestones in the Summer, but both reached 75% total annual growth around the same time. --Consequently, the Summer focused diffuse-porous trees are exposed to higher temperatures and drought risk than ring-porous trees.-- If leaf development is sped up when Spring temperatures are warmer, we should seek to understand whether this relationship holds, and how it potentially varies among trees of different wood-types.

@@

We fit the model of @mcmahon_general_2015 to define *phenological dates and growth rates* (Fig. `r schematic`).

```{r schematic, out.width="100%", echo=FALSE, fig.align="center", fig.cap= schematic_caption}
# Created in RScripts/additional_figures.R
knitr::include_graphics("tables_figures/schematic.png")
```

1. Model was fit to a random sample of 20 out of 106 tags to speed up computation. *(This isn't the case anymore, I fit it to all tag years)*
1. Climwin mean temperatures are recentered at 65F to help sanity check intercepts and slopes. 65F was chosen because it is a "typical" climwinmean value *(Changed to 16 C based on updated climwin windows)*
1. The multivariate model for $(DOY_{25}, DOY_{50}, DOY_{75})$ has:
    1. Fixed effects: The interaction between wood type (diffuse vs ring porous) and climwin mean
    1. Random effects: tag number (no species just yet)

@@

*Phenological patterns*

Across all trees, Pooled across all species and all years the average DOY where max growth rate was achieved is DOY: 158, the DOY where 25% was achieved is DOY: 129, 50% - DOY: 157, 75% - DOY:184.

Growth pheonology differed between ring- and diffuse- porous species, with ring-porous species having an earlier start to growth, with growth spread out over a longer growing season (Fig. `r growth_pheno_by_xylem`).
Specifically,

@@
